---
title: "TAB_COMP"
output: html_document
---

Tablet comparison script. 
====

Set up some functions. 

```{r, echo=FALSE}
rm(list=ls())
## for bootstrapping 95% confidence intervals
theta <- function(x,xdata,na.rm=T) {mean(xdata[x],na.rm=na.rm)}
ci.low <- function(x,na.rm=T) {
  mean(x,na.rm=na.rm) - quantile(bootstrap(1:length(x),1000,theta,x,na.rm=na.rm)$thetastar,.025,na.rm=na.rm)}
ci.high <- function(x,na.rm=T) {
  quantile(bootstrap(1:length(x),1000,theta,x,na.rm=na.rm)$thetastar,.975,na.rm=na.rm) - mean(x,na.rm=na.rm)}

na.mean <- function(x) {mean(x,na.rm=T)}
na.sum <- function(x) {sum(x,na.rm=T)}
na.median <- function(x) {median(x,na.rm=T)}
na.cor <- function(x) {cor(x,na.rm=T)}

library(plyr)
library(stringr)
library(ggplot2)
library(lme4)
library(bootstrap)
library(xtable)
theme_set(theme_bw())
```

Load data and preliminaries.

```{r}
tab <- read.csv("~/Projects/tablet_norming/data/tabletstudyresults.csv")
tab$reaction.time <- as.numeric(as.character(tab$reaction.time))
tab$trial.type <- factor(tab$trial.type, c("rec","MEcontrol","MEexperimental"))
```

Now add demographics.

```{r}
demo <- read.csv("~/Projects/tablet_norming/data/tablet_demographics.csv")
demo$age.group <- floor(demo$age)
sum(demo$exclude)
aggregate(subject.id ~ exclusion.crit, demo, length)
```

Merge, remove exclusions and fillers, and look at age distribution.

```{r}
d <- merge(tab, demo)
d <- subset(d, exclude == 0)
d <- subset(d, trial.type != "filler")
```

Reaction time distribution. Based on this plot, we prune at +/- 2SDs in log space (red lines). We also remove incorrect trials.

```{r}
d$reaction.time[d$response=="N"] <- NA # start by removing 

m <- mean(log(d$reaction.time)) 
s <- sd(log(d$reaction.time))

qplot(reaction.time/1000,
      data=d) + 
  geom_vline(xintercept=exp(m - 2*s)/1000, col="red",lty=2) + 
  geom_vline(xintercept=exp(m + 2*s)/1000, col="red",lty=2) + 
  scale_x_log10(breaks=c(1,2,5,10,20,50))

mean(d$reaction.time)
median(d$reaction.time)
exp(m-(2*s))
exp(m+(2*s))
mean(d$reaction.time > exp(m + 2*s) |
       d$reaction.time < exp(m - 2*s))
```

Do the pruning here.

```{r}
d$reaction.time[d$reaction.time > exp(m + 2*s) |
                    d$reaction.time < exp(m - 2*s)] <- NA
```


How many participants do we have? 

```{r}
print(length(unique(d$subject.id)))

dems <- ddply(d, .(age.group,subject.id), summarise,
              age = age[1],
              male = gender[1] == "M")
dems <- ddply(dems, .(age.group), summarise,
      n = length(subject.id),
      m.age = mean(age),
      m.male = mean(male))
xtable(dems)
```

Did they finish?

```{r}
d$trial.number <- as.numeric(as.character(d$trial.number))
mst <- ddply(d, .(subject.id,age.group), summarise,
             max.trial=max(trial.number))

msta <- ddply(mst, .(age.group), summarise,
              finished = mean(max.trial==28), 
              num.trials = mean(max.trial))
xtable(msta)
```


Actual analysis starts here
====

Aggregation. 

```{r}
mss <- ddply(d, .(subject.id, trial.type, age.group), summarise,
             RT=mean(reaction.time, na.rm=TRUE), 
             correct=sum(response=="Y") / sum(response=="Y" | response=="N"))
```

Accuracy:

```{r}
msa <- ddply(mss, .(trial.type, age.group), summarise, 
             rt=mean(RT,na.rm=TRUE), 
             rt.cih=ci.high(RT),
             rt.cil=ci.low(RT),
             acc=mean(correct,na.rm=TRUE),
             acc.cih=ci.high(correct),
             acc.cil=ci.low(correct)) 

levels(msa$trial.type) <- c("Familiar Word","ME Control","ME Inference")
# pdf("~/Projects/dlet_norming/writeup/figures/accuracy.pdf",width=5,height=3)
qplot(age.group, acc,
      fill=trial.type,
      ymin=acc-acc.cil,
      ymax=acc+acc.cih,
      position=position_dodge(width=.9), 
      geom=c("bar","linerange"),
      stat="identity",
      data=msa) + 
  xlab("Age (Years)") + 
  ylab("Accuracy") + 
  scale_fill_discrete(name="Trial Type") + 
  geom_hline(yintercept=.5,lty=2)
# dev.off()
```

and RT:

```{r}
# pdf("~/Projects/dlet_norming/writeup/figures/rt.pdf",width=5,height=3)
qplot(age.group, rt/1000,
      fill=trial.type,
      ymin=rt/1000-rt.cil/1000,
      ymax=rt/1000+rt.cih/1000,
      position=position_dodge(width=.9), 
      geom=c("bar","linerange"),
      stat="identity",
      data=msa) + 
  xlab("Age (Years)") + 
  ylab("Reaction Time (s)") + 
  scale_fill_discrete(name="Trial Type") 
# dev.off()

```

Continuous age analysis
-----

Accuracy continuous:

```{r}
mss <- ddply(subset(d,trial.type!="MEexperimental"), 
             .(subject.id, age), summarise,
             rt=mean(reaction.time, na.rm=TRUE), 
             correct=sum(response=="Y") / sum(response=="Y" | response=="N"))

qplot(age,correct, 
      data=mss) + 
  xlim(c(1,5)) + 
  scale_y_continuous(breaks=seq(0,1,.25), limits=c(0,1.1)) + 
  geom_hline(yintercept=.5,lty=2) +
  geom_smooth()
```

Now with all trial types. (Consolidate rec and ME control). 

```{r}
d$trial.type.simple <- d$trial.type
levels(d$trial.type.simple) <- c("Familiar","Familiar","ME")
mss <- ddply(d, 
             .(subject.id, trial.type.simple, age), summarise,
             rt=mean(reaction.time, na.rm=TRUE), 
             rt.cih=ci.high(reaction.time),
             rt.cil=ci.low(reaction.time),
             correct=sum(response=="Y") / sum(response=="Y" | response=="N"))

qplot(age,correct, col=trial.type.simple, 
      data=mss) + 
  scale_y_continuous(breaks=seq(0,1,.25), limits=c(0,1.1)) + 
  geom_hline(yintercept=.5,lty=2) +
  geom_smooth()
```

and RT

```{r}
# pdf("~/Projects/dlet_norming/writeup/figures/individuals.pdf",width=5,height=3)
qplot(age, rt/1000, geom="pointrange",
      ymin=rt/1000-rt.cil/1000,
      ymax=rt/1000+rt.cih/1000,
      data=subset(mss, trial.type.simple="Familiar")) + 
  ylim(c(0,7)) + 
  ylab("Reaction Time (s)") +
  xlab("Age Group (Years)") + 
  geom_smooth()
# dev.off()
```

Item group analysis
-----

Accuracy:

```{r}
d$easy.item <- d$word %in% c("dog","cat","cookie","bottle",
                                 "cup","car","shoe","apple")
mss <- ddply(subset(d,trial.type!="MEexperimental"), 
             .(subject.id, trial.type, easy.item, age.group), summarise,
             RT=mean(reaction.time, na.rm=TRUE), 
             correct=sum(response=="Y") / sum(response=="Y" | response=="N"))

msa <- ddply(mss, .(age.group, easy.item), summarise, 
             rt=mean(RT,na.rm=TRUE), 
             rt.cih=ci.high(RT),
             rt.cil=ci.low(RT),
             acc=mean(correct,na.rm=TRUE),
             acc.cih=ci.high(correct),
             acc.cil=ci.low(correct)) 

qplot(age.group, acc,
      fill=easy.item,
      ymin=acc-acc.cil,
      ymax=acc+acc.cih,
      position=position_dodge(width=.9), 
      geom=c("bar","linerange"),
      stat="identity",
      data=msa) + 
  geom_hline(yintercept=.5,lty=2)
```

and RT:

```{r}
qplot(age.group, rt,
      fill=easy.item,
      ymin=rt-rt.cil,
      ymax=rt+rt.cih,
      position=position_dodge(width=.9), 
      geom=c("bar","linerange"),
      stat="identity",
      data=msa)
```

Item analysis
----

```{r}
mssi <- ddply(d, .(subject.id, trial.type, word), summarise,
              RT=mean(reaction.time, na.rm=TRUE), 
              correct=sum(response=="Y") / sum(response=="Y" | response=="N"))

msi <- ddply(mssi, .(trial.type, word), summarise, 
             rt=mean(RT,na.rm=TRUE), 
             rt.cih=ci.high(RT),
             rt.cil=ci.low(RT),
             acc=mean(correct),
             acc.cih=ci.high(correct),
             acc.cil=ci.low(correct)) 

msi$word <- reorder(msi$word, msi$acc)

qplot(word, acc,
      ymin=acc-acc.cil,
      ymax=acc+acc.cih,
      geom="pointrange",
      data=msi) + 
  facet_wrap(~trial.type,scales="free_x") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=.5))
```

and RT:

```{r}
msi$word <- reorder(msi$word, msi$rt)

qplot(word, rt,
      ymin=rt-rt.cil,
      ymax=rt+rt.cih,
      geom="pointrange",
      data=msi) + 
  facet_wrap(~trial.type,scales="free_x") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=.5))
```

Number of trials
----

Nearly everyone sticks it out to the end!

```{r}
d$trial.number <- as.numeric(as.character(d$trial.number))
mst <- ddply(d, .(subject.id,age.group), summarise,
             max.trial=max(trial.number))

qplot(max.trial, facets=~age.group, data=mst)
```

and means across groups.

```{r}
mstm <- ddply(mst, .(age.group), summarise, 
              trials = mean(max.trial), 
              trials.cih = ci.high(max.trial), 
              trials.cil = ci.low(max.trial))
qplot(age.group, trials,
      fill=factor(age.group),
      ymin=trials-trials.cil,
      ymax=trials+trials.cih,
      position=position_dodge(width=.9), 
      geom=c("bar","linerange"),
      stat="identity",
      data=mstm) + 
  geom_hline(yintercept=28, lty=2)
```

and RT means across trials:

```{r}
mst <- ddply(d, .(trial.number, age.group), summarise,
             RT=mean(reaction.time, na.rm=TRUE))

qplot(trial.number, RT, data=mst) + 
  geom_smooth(method="lm")
```

Statisics 
----

```{r}
library(xtable)
rt.mod <- lmer(reaction.time ~ trial.number + age * trial.type + 
                 (trial.type | subject.id), 
               data=d)
summary(rt.mod)
xtable(summary(rt.mod)$coef)

d$acc <- d$response=="Y"
acc.mod <- lmer(acc ~ age.group * trial.type + 
                  (trial.type | subject.id), 
                family="binomial",
                data=d)
summary(acc.mod)
xtable(summary(acc.mod)$coef)

```

Reliability
----

Fairest analysis -- split halves where 1 half is the odd-numbered Familiar trials and even numbered ME trials, and the other half the remaining trials Control trials.

```{r}
library(dplyr)
bytrial <- d %>%
  select(age.group,subject.id,trial.type,trial.number,acc,reaction.time) %>%
  group_by(age.group,trial.type,subject.id) %>%
  filter(trial.type != "MEexperimental") %>%
  arrange(trial.number) %>%
  mutate(trial.order = 1:length(trial.number),
         even.half = trial.order %% 2 == 0,
         first.half = (even.half & trial.type == "rec" ) | 
           (!even.half & trial.type != "rec" ))

type.trials <- bytrial %>%
  group_by(age.group,subject.id,trial.type,add=FALSE) %>%
  summarise(acc = mean(acc),
            rt = na.median(reaction.time))
type.trials <- reshape(type.trials, idvar = c("subject.id","age.group"),
                       timevar = "trial.type", direction="wide")

half.trials <- bytrial %>%
  group_by(age.group,subject.id,first.half,add=FALSE) %>%
  summarise(acc = mean(acc),
            rt = na.median(reaction.time))
half.trials <- reshape(half.trials, idvar = c("subject.id","age.group"),
                       timevar = "first.half", direction="wide")

```

Statistics

``` {r}
reliability.half <- half.trials %>%
  group_by(age.group) %>%
  summarise(acc.corr = cor(acc.TRUE,acc.FALSE,use="complete.obs"),
            acc.corr.t = cor.test(acc.TRUE,acc.FALSE,
                                  use="complete.obs")$statistic,
            acc.corr.p = cor.test(acc.TRUE,acc.FALSE,
                                  use="complete.obs")$p.value,
            rt.corr = cor(rt.TRUE,rt.FALSE,use="complete.obs"),
            rt.corr.t = cor.test(rt.TRUE,rt.FALSE,
                                 use="complete.obs")$statistic,
            rt.corr.p = cor.test(rt.TRUE,rt.FALSE,
                                 use="complete.obs")$p.value)

reliability.type <- type.trials %>%
  group_by(age.group) %>%
  summarise(acc.corr = cor(acc.rec,acc.MEcontrol,use="complete.obs"),
            acc.corr.t = cor.test(acc.rec,acc.MEcontrol,
                                  use="complete.obs")$statistic,
            acc.corr.p = cor.test(acc.rec,acc.MEcontrol,
                                  use="complete.obs")$p.value,
            rt.corr = cor(rt.rec,rt.MEcontrol,use="complete.obs"),
            rt.corr.t = cor.test(rt.rec,rt.MEcontrol,
                                 use="complete.obs")$statistic,
            rt.corr.p = cor.test(rt.rec,rt.MEcontrol,
                                 use="complete.obs")$p.value)

print(reliability.half)
print(reliability.type)
```

Spearman Brown Prophecy Formula

``` {r}
sb.prophecy <- function(m,r) {
  (m * r) / (1 + (m-1) * r)
  }

reliability.half <- mutate(reliability.half, 
                           acc.corr.sb = sb.prophecy(2, acc.corr),
                           rt.corr.sb = sb.prophecy(2, rt.corr))
reliability.type <- mutate(reliability.type, 
                           acc.corr.sb = sb.prophecy(2, acc.corr),
                           rt.corr.sb = sb.prophecy(2, rt.corr))
```

Plots

```{r}
qplot(rt.FALSE,rt.TRUE,
      col=factor(age.group),
      position="jitter",
      data=half.trials) + 
  geom_smooth(method="lm", se=FALSE) + 
  geom_abline(intercept=0,slope=1,lty=2)
```