---
title: "TAB_COMP"
output: html_document
---

Tablet comparison script. 
====

Set up some functions. 

```{r}
rm(list=ls())
source("~/Projects/R/Ranalysis/useful_dplyr.R")
library(xtable)
library(psych)
```

Reading in experimental data
------

Demographics. 
```{r}
tab.demo <- read.csv("../data/tablet_demographics.csv")
tab.demo <- tab.demo %>% rename(subid = SID)

story.demo <- read.csv("../data/story_demographics.csv")
story.demo <- story.demo %>% rename(subid = SID)

et.demo <- read.csv("../data/et_demographics.csv")
et.demo <- et.demo %>% rename(subid = SID)
```

First read tablet.

```{r}
tab <- read.csv("../data/tabletstudyresults.csv")
tab$reaction.time <- as.numeric(as.character(tab$reaction.time))
tab$trial.type <- factor(tab$trial.type, 
                         levels = c("rec","MEcontrol","MEexperimental","filler"),
                         labels = c("FAM-fam","FAM-nov","NOV-fam","filler"))
```

Then read storybook.

```{r}
story <- read.csv("../data/storystudyresults.csv")
story <- story %>% 
  gather(trial, correct, ends_with(".Correct"), na.rm=FALSE) %>%
  select(-starts_with("X")) %>%
  mutate(trial = as.numeric(str_replace_all(str_replace(trial, "X", ""), ".Correct", "")))
lists <- read.csv("../data/lists.csv")

story <- left_join(story, lists, by=c("List","trial"))

story$trial.type <- factor(story$trial.type, 
                         levels = c("familiar-familiar","FAMILIAR-novel",
                                    "familiar-NOVEL","filler"),
                         labels = c("FAM-fam","FAM-nov","NOV-fam", "filler"))

```

Then read eye-tracking.

```{r}
et <- read.csv("../eye_tracking/R/eye.tracking.csv", row.names="X")
et$trial.type <- factor(et$word.type, 
                         levels = c("Familiar-Familiar","Familiar-Novel",
                                    "Novel-Familiar"),
                         labels = c("FAM-fam","FAM-nov","NOV-fam"))
```

Merge these three together. This is quite ugly because of differing naming conventions across data files.

```{r}
et <- et %>% 
  select(sid, whole.trial.num, trial.type, rt, prop) %>%
  rename(subid = sid, 
         trial.num = whole.trial.num,
         correct = prop)

et <- left_join(et, et.demo, by="subid")
et$condition  <- "eye-tracker"
et$rt <- et$rt * 1000 # convert to ms

tab <- tab %>%
  mutate(correct = response=="Y") %>%
  select(subject.id, trial.number, trial.type, reaction.time, correct) %>%
  rename(subid = subject.id, 
         trial.num = trial.number, 
         rt = reaction.time)

tab <- left_join(tab, tab.demo)
tab$condition  <- "tablet"

story <- story %>%
  select(SID, trial, trial.type, correct) %>%
  rename(subid = SID, 
         trial.num = trial)

story <- left_join(story, story.demo)
story$condition <- "storybook"

d <- rbind_list(et, tab, story) %>%
  mutate(age.grp = floor(age))
```

Demographics and filtering
--------------------------

Report function.

```{r}
report <- function(x) {
x %>%
  group_by(subid, condition, age.grp) %>%
  summarise(age = age[1],
            sex = sex[1]) %>% 
  group_by(condition, age.grp, add=FALSE) %>% 
  summarise(n=n(), 
            m.age = mean(age),
            m.male = mean(sex=="male"))
}
```

Filter for ages and filler trials. 

```{r}
d <- d %>% 
  filter(age.grp > 0, age.grp < 5, !is.na(age.grp))     
d <- d %>% filter(trial.type != "filler")
```

Non-inclusion criteria that don't relate to methods.

```{r}
d %>% group_by(subid, condition, exclusion.crit) %>%
  summarise(n = 1) %>%
  group_by(condition, exclusion.crit, add=FALSE) %>%
  summarise(n=sum(n)) 

n <- d %>% 
  filter(exclusion.crit != "lang", 
         exclusion.crit != "dd")
```

Now, other exclusion criteria.

```{r}
d %>% 
  group_by(subid, condition, exclusion.crit) %>%
  summarise(n = 1) %>%
  group_by(condition, exclusion.crit, add=FALSE) %>%
  summarise(n= sum(n)) %>%
  group_by(condition, add=FALSE) %>% 
  mutate(prop = n/ sum(n))

d <- d %>% 
  filter(exclude == 0)

xtable(report(d))

```

Data-based exclusions
---------------------

Did they finish?

```{r}
qplot(trial.num, facets=~condition, data=d)

mst <- d %>% group_by(subid, condition, age.grp) %>%
  summarise(max.trial = max(trial.num)) %>%
  group_by(condition, age.grp, add=FALSE) %>%
  summarise(finished = ifelse(condition[1]=="storybook", mean(max.trial==23), 
                              mean(max.trial==28)), 
            num.trials = mean(max.trial))

qplot(age.grp, finished, col=condition, 
      geom="line", stat="identity",
      position="dodge",
      data=mst)
```


Reaction time pruning
---------------------

Reaction time distribution.  We also remove incorrect trials for the tablet, and target-initial trials for the eye-tracker.

```{r}
d$rt[d$condition == "tablet" & !d$correct] <- NA # start by removing false trials
d$rt[d$condition == "eye-tracker" & d$rt < 317] <- NA # remove T initial trials

d %>% 
  group_by(condition) %>%
  summarise(m = mean(log(rt), na.rm=TRUE), 
            s = sd(log(rt), na.rm=TRUE))

qplot(rt/1000,
      data=filter(d, condition != "storybook")) + 
  facet_grid(.~condition) + 
#   geom_vline(xintercept=exp(m - 2*s)/1000, col="red",lty=2) + 
#   geom_vline(xintercept=exp(m + 2*s)/1000, col="red",lty=2) + 
  scale_x_log10(breaks=c(1,2,5,10,20,50))

```

Age effects:

```{r}
ggplot(d) + 
  geom_bar(aes(x = rt/1000, y=..count..)) +
    scale_x_log10(breaks=c(1,5,10,50)) +
  facet_grid(condition~age.grp)
```

Prune excessive reaction times. (Arbitrary threshold). 

```{r}
d$reaction.time[d$rt > 4000 |
                    d$rt < 500] <- NA

```


Actual analysis starts here
====

Aggregation. 

```{r}
mss <- d %>%
  group_by(subid, trial.type, age.grp, condition) %>%
  summarise(age=age[1],
            rt=mean(rt, na.rm=TRUE), 
            correct=mean(correct)) 

mss$trial.type <- factor(mss$trial.type, 
                        levels = c("FAM-fam","FAM-nov","NOV-fam"), 
                        labels=c("Familiar Word","ME Control","ME Inference"))

ms <- mss %>%
  group_by(trial.type, age.grp, condition, add=FALSE) %>%
  summarise(rt.cih=ci.high(rt),
            rt.cil=ci.low(rt),
            rt=mean(rt,na.rm=TRUE), 
            acc.cih=ci.high(correct),
            acc.cil=ci.low(correct),
            acc=mean(correct,na.rm=TRUE)) 
```

Accuracy as a barplot.

```{r}
pdf("~/Projects/tablet_norming/writeup/figures/accuracy.pdf",width=7,height=5)
qplot(age.grp, acc,
      fill=trial.type,
      ymin=acc-acc.cil,
      ymax=acc+acc.cih,
      position=position_dodge(width=.9), 
      geom=c("bar","linerange"),
      stat="identity",
      data=ms) + 
  facet_grid(condition ~ .) + 
  xlab("Age (Years)") + 
  ylab("Accuracy") + 
  geom_hline(yintercept=.5,lty=2) + 
  scale_fill_discrete(name="Trial Type") 

dev.off()
```

Accuracy (continuous)

```{r}
qplot(age, correct, col=trial.type, data=mss) + 
  facet_grid(. ~ condition) + 
  geom_smooth(method="lm", formula=y ~ inv.logit(x)) + 
  geom_hline(yintercept=.5,lty=2) + 
  ylim(c(0,1.2))
```

and RT:

```{r}
pdf("~/Projects/tablet_norming/writeup/figures/rt.pdf",width=7,height=4)
qplot(age.grp, rt/1000,
      fill=trial.type,
      ymin=rt/1000-rt.cil/1000,
      ymax=rt/1000+rt.cih/1000,
      position=position_dodge(width=.9), 
      geom=c("bar","linerange"),
      stat="identity",
      data=filter(ms, condition != "storybook")) + 
  xlab("Age (Years)") + 
  ylab("Reaction Time (s)") + 
  scale_fill_discrete(name="Trial Type") + 
  facet_grid(condition ~ . , scales="free_y") 
dev.off()
```

Number of trials
----

Nearly everyone sticks it out to the end!

```{r}
mst <- d %>% 
  group_by(subid, age.grp, condition) %>%
  summarise(max.trial = max(trial.num))

qplot(max.trial, facets=condition~age.grp, data=mst)
```

and means across groups.

```{r}
mstm <-  mst %>% 
  group_by(age.grp, condition, add=FALSE) %>%
  summarise(trials = mean(max.trial), 
            trials.cih = ci.high(max.trial), 
            trials.cil = ci.low(max.trial))

qplot(age.grp, trials, 
      fill=factor(age.grp),
      ymin=trials-trials.cil,
      ymax=trials+trials.cih,
      facets=.~condition, 
      position=position_dodge(width=.9), 
      geom=c("bar","linerange"),
      stat="identity",
      data=mstm) + 
  geom_hline(yintercept=28, lty=2)
```

Statistics 
----

t-tests vs. chance.

Accuracies for eye-tracker. 

```{r}
t.test(mss$correct[mss$condition == "eye-tracker" & 
                     mss$trial.type == "Familiar Word" & 
                     mss$age.grp == 1] - .5)
t.test(mss$correct[mss$condition == "eye-tracker" & 
                     mss$trial.type == "ME Control" & 
                     mss$age.grp == 1] - .5)
t.test(mss$correct[mss$condition == "eye-tracker" & 
                     mss$trial.type == "ME Inference" & 
                     mss$age.grp == 1] - .5)
t.test(mss$correct[mss$condition == "eye-tracker" & 
                     mss$trial.type == "ME Inference" & 
                     mss$age.grp == 2] - .5)
```

Other accuracies.

```{r}
t.test(mss$correct[mss$condition == "storybook" & 
                     mss$trial.type == "Familiar Word" & 
                     mss$age.grp == 1] - .5)
t.test(mss$correct[mss$condition == "tablet" & 
                     mss$trial.type == "Familiar Word" & 
                     mss$age.grp == 1] - .5)
t.test(mss$correct[mss$condition == "tablet" & 
                     mss$trial.type == "ME Control" & 
                     mss$age.grp == 1] - .5)
```

Accuracy model. 

```{r}
acc.mod <- glmer(correct ~ age * trial.type * condition + 
                  (trial.type | subid), 
                family="binomial",
                data=filter(d, condition != "eye-tracker"))
summary(acc.mod)
```

RT models. 

```{r}
rt.mod <- lmer(log(rt) ~ age * trial.type * condition + 
                 (trial.type | subid), 
               data=filter(d, condition != "storybook"))
summary(rt.mod)

et.rt.mod <- lmer(log(rt) ~ age * trial.type + 
                 (trial.type | subid), 
               data=filter(d, condition == "eye-tracker"))
summary(et.rt.mod)

tab.rt.mod <- lmer(log(rt) ~ age * trial.type + 
                 (trial.type | subid), 
               data=filter(d, condition == "tablet"))
summary(tab.rt.mod)


```

Reliability
----

We compute crohnbach's alpha. This works well, as long as you do a lot of filtering of columns that have null observations in them (this is what the `get.frame` function does). 

```{r}
get.frame <- function(x) {
  x <- as.data.frame(x)
  non.null.cols <- apply(x, 2, function(x) {!all(is.na(x))})
  x <- x[, non.null.cols]
  return(x)
}
  
acc.alphas <- d %>%
  filter(trial.type != "NOV-fam", !is.na(correct)) %>%
  select(subid, age.grp,  trial.num, condition, correct) %>%
  group_by(age.grp, condition, subid) %>%
  arrange(subid, trial.num) %>%
  mutate(trial.order = str_c("t",as.character(1:length(trial.num))), 
         n = length(trial.num)) %>%
  select(-trial.num) %>%
  spread(trial.order, correct) %>%
  group_by(age.grp, condition,  add=FALSE) %>%
  summarise(n = mean(n), 
    raw.alpha = alpha(get.frame(cbind(t1, t2, t3, t4, t5, t6, t7, t8, 
                                    t9, t10, t11, t12, t13, t14, 
                                    t15, t16)), 
                          na.rm=TRUE, delete=TRUE, 
                          check.keys=FALSE)$total$raw_alpha,
            std.alpha = alpha(get.frame(cbind(t1, t2, t3, t4, t5, t6, t7, t8, 
                                    t9, t10, t11, t12, t13, t14, 
                                    t15, t16)), 
                          na.rm=TRUE, delete=TRUE, 
                          check.keys=FALSE)$total$std.alpha)

## no ME trials, no NA rts, no storbook (no rts collected)
rt.alphas <- d %>%
  filter(trial.type != "NOV-fam", !is.na(rt), 
         condition != "storybook") %>% 
  select(subid, age.grp,  trial.num, condition, rt) %>%
  group_by(age.grp, condition, subid) %>%
  arrange(subid, trial.num) %>%
  mutate(trial.order = str_c("t",as.character(1:length(trial.num))), 
         n = length(trial.num)) %>%
  select(-trial.num) %>%
  spread(trial.order, rt) %>%
  group_by(age.grp, condition,  add=FALSE) %>%
  summarise(n = mean(n),
            alpha = alpha(get.frame(cbind(t1, t2, t3, t4, t5, t6, t7, t8, 
                                          t9, t10, t11, t12, t13, t14, 
                                          t15, t16)), 
                          na.rm=TRUE, delete=TRUE, 
                          check.keys=FALSE)$total$raw_alpha, 
            std.alpha = alpha(get.frame(cbind(t1, t2, t3, t4, t5, t6, t7, t8, 
                                    t9, t10, t11, t12, t13, t14, 
                                    t15, t16)), 
                          na.rm=TRUE, delete=TRUE, 
                          check.keys=FALSE)$total$std.alpha)

xtable(acc.alphas)
xtable(rt.alphas)
```


Much prior reliability work
---------------------------

This split half thing is very unstable...


```{r}
bytrial <- d %>%
  select(subid, age.grp, trial.type, trial.num, condition, correct, rt) %>%
  group_by(age.grp, subid, condition) %>%
  filter(trial.type != "NOV-fam") %>%
  arrange(trial.num) %>%
  mutate(trial.order = 1:length(trial.num),
         even.half = factor(c("odd","even")
                              [as.numeric(trial.order %% 2 == 0)+1]),
         n = length(trial.num)) 

#          first.half = (even.half & trial.type == "rec" ) | 
#            (!even.half & trial.type != "rec" ))

half.trials.acc <- bytrial %>%
  filter(n > 12) %>%
  select(-rt) %>%
  group_by(age.grp, subid, even.half, condition, add=FALSE) %>%
  summarise(acc = mean(correct, na.rm=TRUE)) %>%
  spread(even.half, acc)

half.trials.rt <- bytrial %>%
  filter(n>4) %>%
  select(-correct) %>%
  group_by(age.grp, subid, even.half, condition, add=FALSE) %>%
  summarise(rt = mean(rt, na.rm=TRUE)) %>%
  spread(even.half, rt)
```

Statistics

``` {r}

sb.prophecy <- function(m,r) {
  (m * abs(r)) / (1 + (m-1) * abs(r))
  }

reliability.half <- 
  left_join(half.trials.acc %>%
              group_by(age.grp, condition) %>%
              summarise(acc.corr = cor(even, odd, 
                                       use="complete.obs")) %>%
              mutate(acc.corr.sb = sb.prophecy(2, acc.corr)), 
            half.trials.rt %>%
              filter(condition != "storybook") %>%
              group_by(age.grp, condition) %>%
              summarise(rt.corr = cor(even, odd, 
                                       use="complete.obs")) %>%
              mutate(rt.corr.sb = sb.prophecy(2, rt.corr)))
            
#             acc.corr.t = cor.test(even, odd,
#                                   use="complete.obs")$statistic,
#             acc.corr.p = cor.test(even, odd,
#                                   use="complete.obs")$p.value)

print(reliability.half)
```


```{r}
n.samps <- 1000
sims <- data.frame()
for (i in 1:n.samps) {  
  split <- sample(1:16,8)
    
  bytrial <- d %>%
    select(subid, age.grp, trial.type, trial.num, condition, correct, rt) %>%
    group_by(age.grp, subid, condition) %>%
    filter(trial.type != "NOV-fam") %>%
    arrange(trial.num) %>%
    mutate(trial.num.seq = 1:length(trial.num),
           first.half = c("first","second")[as.numeric(trial.num.seq %in% split)+1], 
           n = length(trial.num.seq))
  
  half.trials.acc <- bytrial %>%
    filter(n > 6) %>% ## ONLY KEEP MORE THAN 6 TRIALS
    select(-rt) %>%
    group_by(subid) %>%
    sample_frac(1, replace=TRUE) %>% ## RESAMPLE OVER PARTICIPANTS
    group_by(age.grp, subid, first.half, condition, add=FALSE) %>%
    summarise(acc = mean(correct, na.rm=TRUE)) %>%
    spread(first.half, acc)
  
  half.trials.rt <- bytrial %>%
    filter(!is.na(rt)) %>%
    mutate(n = length(rt)) %>%
    filter(n > 4) %>% ## ONLY KEEP MORE THAN 4 TRIALS
    select(-correct) %>%
    group_by(subid) %>% ## RESAMPLE OVER PARTICIPANTS
    sample_frac(1, replace=TRUE) %>%
    group_by(age.grp, subid, first.half, condition, add=FALSE) %>%
    summarise(rt = mean(rt, na.rm=TRUE)) %>%
    spread(first.half, rt)
  
  reliability.half <- 
    left_join(half.trials.acc %>%
                group_by(age.grp, condition) %>%
                summarise(acc.corr = cor(first, second, 
                                         use="complete.obs")) %>%
                mutate(acc.corr.sb = sb.prophecy(2, acc.corr)), 
              half.trials.rt %>%
                filter(condition != "storybook") %>%
                group_by(age.grp, condition) %>%
                summarise(rt.corr = cor(first, second, 
                                        use="complete.obs")) %>%
                mutate(rt.corr.sb = sb.prophecy(2, rt.corr))) %>%
    mutate(sim = i)
  
  sims <- bind_rows(sims, reliability.half)
}  

ms <- sims %>% 
  group_by(age.grp, condition) %>%
  summarise(m.acc.corr = mean(acc.corr, na.rm=TRUE),
            m.rt.corr = mean(rt.corr, na.rm=TRUE),
            m.acc.sb = mean(acc.corr.sb, na.rm=TRUE),
            m.rt.sb = mean(rt.corr.sb, na.rm=TRUE),
            cil.acc.corr = quantile(acc.corr, c(.025), na.rm=TRUE),
            cih.acc.corr = quantile(acc.corr, c(.975), na.rm=TRUE),
            cil.rt.corr = quantile(rt.corr, c(.025), na.rm=TRUE),
            cih.rt.corr = quantile(rt.corr, c(.975), na.rm=TRUE),
            cil.acc.sb = quantile(acc.corr.sb, c(.025), na.rm=TRUE),
            cih.acc.sb = quantile(acc.corr.sb, c(.975), na.rm=TRUE),
            cil.rt.sb = quantile(rt.corr.sb, c(.025), na.rm=TRUE),
            cih.rt.sb = quantile(rt.corr.sb, c(.975), na.rm=TRUE)) %>%
  gather(measure, value, starts_with("m."), starts_with("ci")) %>%
  separate(measure, c("stat","dv", "method"), sep="\\.") %>%
  spread(stat, value)
```

Plot this

```{r}
qplot(age.grp, m,
      ymin = cil,
      ymax = cih,
      facets = . ~ dv, 
      position=position_dodge(width=.05),
      col=condition, geom=c("line","linerange"),
      data=filter(ms, method=="corr")) + 
  ylim(c(-.25,1)) + 
  ylab("Split-Half Reliability (Average Pearson r)") + 
  xlab("Age Group")
```

Individual differences
----------------------

```{r}
d$trials.consol <- revalue(factor(d$trial.type), 
                           c("FAM-fam"="familiar",
                             "FAM-nov"="familiar",
                             "NOV-fam"="me"))
rs.rt <- d %>%
  group_by(subid, trials.consol, age.grp, condition) %>%
  summarise(rt=mean(rt, na.rm=TRUE), 
            correct=mean(correct)) %>%
  group_by(condition, age.grp, subid) %>%
  select(-correct) %>%
  spread(trials.consol, rt) %>%
  group_by(condition, age.grp) %>%
  summarise(r = cor(familiar, me, use="pairwise.complete.obs"))
            
rs.acc <- d %>%
  group_by(subid, trials.consol, age.grp, condition) %>%
  summarise(rt=mean(rt, na.rm=TRUE), 
            correct=mean(correct)) %>%
  group_by(condition, age.grp, subid) %>%
  select(-rt) %>%
  spread(trials.consol, correct) %>%
  group_by(condition, age.grp) %>%
  summarise(r = cor(familiar, me, use="pairwise.complete.obs"))
s
```

